Model Search in Relational Domains

### 2016-10-23 domains.tex
section 3 updates

% \paragraph{Interval Abstract Domain} 
% The {\em Interval} abstract domain is a lattice 
% $ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
% $ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
% the set of intervals of type $[l,u]$ over numeric data 
% type with $l \leq u$. The least element is $\bot$ and the 
% greatest element is $\top$ which maps all variables to their
% minimum ($min$) and maximum ($max$) values.  An interval 
% $\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
% partial order $\sle_{I}$ over elements in the set $Itv$ is 
% given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
% A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
% $\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
% A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
% $\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
% The galois connection between the concrete environment and intervals is 
% given as follows.
% \[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
% \[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
%
\begin{wrapfigure}{r}{4.5cm}
\vspace*{-7ex}
\scalebox{.65}{\import{figures/}{octagons.pspdftex}}
\caption{Example of an octagon}\label{octagon}
\vspace*{-5ex}
\end{wrapfigure} 
%
In the interval domain $\intervals[\numvars]$,
for a program with $N=|\numvars|$ variables, the matrix $\mat{C}$ 
has $2N$ rows. Hence, it generates at most $2N$ inequalities, one
for the upper and lower bounds of each variable.
%
For octagons $\octagons[\numvars]$, we have at most $2N^2$
inequalities, one for the upper and lower bounds of each variable and
sums and differences for each pair of variables. Figure~\ref{octagon}
gives an example of an octagon and its associated inequalities.
\Omit{
The interval domain is a non-relational domain because a single
inequality only contains a single variable.  The octagon domain, 
however, is relational.  
}
%
A fundamental difference between relational and 
non-relational domains is that the relational
domains require the computation of a \emph{closure} 
in order to obtain a normal form that is usually required for precise 
%and efficient 
domain operations.  An example of a closure computation for octagonal 
inequalities is given by
$\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z
\leq 5) \wedge (x-z \leq 9))$.
%below.
%\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
%
The closure computes all implied domain constraints.  For octagons,
the closure operation is very expensive (cubic complexity in the
number of program variables).
%
%We will see that our model search algorithm performs closure computations lazily,
%postponing them to the point where they are actually necessary.\rmcmt{are we
%showing this ?}

>>>>

% \paragraph{Interval Abstract Domain} 
% The {\em Interval} abstract domain is a lattice 
% $ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
% $ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
% the set of intervals of type $[l,u]$ over numeric data 
% type with $l \leq u$. The least element is $\bot$ and the 
% greatest element is $\top$ which maps all variables to their
% minimum ($min$) and maximum ($max$) values.  An interval 
% $\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
% partial order $\sle_{I}$ over elements in the set $Itv$ is 
% given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
% A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
% $\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
% A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
% $\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
% The galois connection between the concrete environment and intervals is 
% given as follows.
% \[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
% \[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
%
\begin{wrapfigure}{r}{4.5cm}
\vspace*{-7ex}
\scalebox{.65}{\import{figures/}{octagons.pspdftex}}
\caption{Example of an octagon}\label{octagon}
\vspace*{-5ex}
\end{wrapfigure} 
%
%In the interval domain $\intervals[\numvars]$,
For a program with $N=|\numvars|$ variables, 
the matrix $\mat{C}$ for the interval domain 
$\intervals[\numvars]$, has $2N$ rows. Hence, it 
generates at most $2N$ inequalities, one
for the upper and lower bounds of each variable.
%
For octagons $\octagons[\numvars]$, we have at most $2N^2$
inequalities, one for the upper and lower bounds of each variable and
sums and differences for each pair of variables. Figure~\ref{octagon}
gives an example of an octagon and its associated inequalities.
\Omit{
The interval domain is a non-relational domain because a single
inequality only contains a single variable.  The octagon domain, 
however, is relational.  
}
%
A fundamental difference between relational and 
non-relational domains is that the relational
domains require the computation of a \emph{closure} 
in order to obtain a normal form that is usually required for precise 
%and efficient 
domain operations.  An example of a closure computation for octagonal 
inequalities is given by
$\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z
\leq 5) \wedge (x-z \leq 9))$.
%below.
%\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
%
The closure computes all implied domain constraints.  For octagons,
closure is the most critical and expensive operator (cubic complexity in the
number of program variables)~\cite{pldi15}.
%
%We will see that our model search algorithm performs closure computations lazily,
%postponing them to the point where they are actually necessary.\rmcmt{are we
%showing this ?}
